---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Loading the data

```{r}
data <- read.csv("activity.csv")

head(data)
```

## Calculate the total number of steps taken per day


```{r}
total_Steps_per_day <- aggregate(steps ~ date , data, FUN = sum)

head(total_Steps_per_day)
```

##  histogram of the total number of steps taken each day

```{r}
hist(total_Steps_per_day$steps, xlab = "Total number of steps taken daily") 
```

## Mean and median number of steps taken each day
```{r}
mean_Steps_per_day <- mean(total_Steps_per_day$steps, na.rm = T)
mean_Steps_per_day

median_Steps_per_day <- median(total_Steps_per_day$steps, na.rm = T)
median_Steps_per_day

```

## Time series plot of the average number of steps taken

```{r}
avg_Steps_per_day <- aggregate(steps ~ interval , data, FUN = mean)
plot(avg_Steps_per_day$interval, avg_Steps_per_day$steps, type = "l", main = "Average number of steps taken", xlab = "interval", ylab = "average number of steps taken per day"  )
```

## The 5-minute interval that, on average, contains the maximum number of steps
```{r}
max(avg_Steps_per_day$steps)
```

## strategy for imputing missing data
1. Total number of missing values

```{r}
missing_entries <- sum(is.na(data$steps))
missing_entries
```
2. Replace NA in the dataset with the mean.

```{r}
no_missing_data <- data
no_missing_data$steps [is.na(no_missing_data$steps)] <- mean(data$steps, na.rm = T)
```
3. calculate missing values again after replacing NA entries 
```{r}
no_missing_entries <- sum(is.na(no_missing_data$steps))
no_missing_entries
```

##Histogram of the total number of steps taken each day after missing values are imputed

```{r}
All_Steps_per_day <- aggregate(steps ~ date , no_missing_data, FUN = sum)
head(All_Steps_per_day)
hist(All_Steps_per_day$steps, xlab = "Total number of steps taken daily") 
```

